---
title: "Multi-Years Spawners"
author: "Scott A. Akenhead"
date: today
format: pdf
editor: visual
execute:
  cache: true
  hold: Hold 
---

```{r}
#| label: setup
#| echo: false
#| output: false
library(rstan);library(magrittr);library(ggplot2); 
library(patchwork); library(knitr)
rstan_options(auto_write = TRUE) # do not recompile code not changed

theme_SAA = function(){
  theme_set(theme_bw()) 
  theme_replace(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(
        linetype='dotted',colour ='grey', linewidth=0.2) ) 
}
```

# Intro

**Objective:** Fit a spawner abundance model to one spawning site using many years of data at once. For example, wild Sockeye spawning in the Okanagan River above Osoyoos Lake.

**Model:** Derived from a subset of *Multilevel-Regression-Spawner*: Multiple samples in each of multiple years; one river.

**Simulation:** Derived from *Simulate Spawners*. This version creates observation dates that vary by year.

# Model

This model will predicted the *run* (abundance) each year of spawners from the counts of spawners (observations with sampling error) each year. Runs are simulated from spawner population dynamics including productivity (smolts/spawner), age compostion of smolts and adults, and smolt survival by ocean entry year. Within a year, spawners enter according to annually varying entry time and spread, and stay for a fixed residence time.

> "In general, older and larger wild Atlantic salmon return to freshwater to spawn earlier in the season than younger and smaller salmon, and female salmon return earlier in the season than males (e.g. Foldvik et al., 2024; Harvey et al., 2017; Jokikokko et al., 2004; Niemelä et al., 2006; Shearer, 1990)."\
> – Miettinen, A., Romakkaniemi, A., Dannewitz, J., Pakarinen, T., Palm, S., Persson, L., Östergren, J., Primmer, C.R., and Pritchard, V.L. 2024. Temporal allele frequency changes in large-effect loci reveal potential fishing impacts on salmon life-history diversity. Evolutionary Applications **17**(4): e13690. doi:[10.1111/eva.13690](https://doi.org/10.1111/eva.13690).

This is a multi-level regression (**MLR**), where level 1 parameters vary by year and share information through a distribution fitted to each level 1 parameters. That distribution is the level 2 parameters and the basis for sampling (fitting) the PDDs at level 1. Level 2 is then the overall pattern across years and level 1 is the change from that each year. information is shared between years: trial values for abundance each year are draws from a fitted PDD of abundances (with prior knowledge for that PDD included). That improves the precision of an estimate for a year's run by using information from the other years; estimates for poorly observed years will tend toward the mean run for all years. Similarly for the varying time and spread of **spawner entry** in every year: these parameters are sampled from a PDD fitted to those annual values.

"When data sizes are small, parameter estimation can be stabilized and sharpened by providing **hierarchical priors**. With hierarchical priors, the data determines the amount of *partial pooling* among the groups. The only drawback is that if the number of groups is small, it can be hard to fit these models without strong hyperpriors." "Priors on priors, also known as *hyperpriors*, should be treated the same way as priors on lower-level parameters in that **as much prior information as is available should be brought to bear.** "\
– Stan Development Team. 2024. Stan User’s Guide Version 2.35

For example, the mean (with standard deviation) of spawner run size over all years, *run*, is level 2 and never a negative number. The prior for the distribution for samples of *run* is *lognomal(run_mu, run_sigma)*. The difference from that overall mean in a single year is another mean *run_year* with prior *lognomal(run, run_year_sigma)*. The result of fitting a level estimate of mean run, provides information to the level 1 is also constrained to be non-negative.

## Parameters

The parameters – or the population dynamics underlying annual spawner abundance, for the patterns of spawner abundance within a year, and for the prior knowledge and fitted parameters of a statistical model – could be those of the following table.

+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| Parameter Description                                                                                                                                                                                                                                                                                                                                        | Code\                   | Math\                     |
|                                                                                                                                                                                                                                                                                                                                                              | Variables               | Symbol                    |
+==============================================================================================================================================================================================================================================================================================================================================================+=========================+===========================+
| **Run**. *Multi-level.* Abundance, total number of spawners. Multiple ages lumped. There is a mean (mu) and stdev (sigma) for each. Runs are simulated, but four starting values are required:\                                                                                                                                                              | Simulation:\            | $r_\mu,r_\sigma$,\        |
| *Trunc_rnorm(n=4, run_mu, run_sigma, lowest=1000).\                                                                                                                                                                                                                                                                                                          | run_init_mu\            | $r_{\mu,y},r_{\sigma, y}$ |
| *The year effect for run size is centered, prior:\                                                                                                                                                                                                                                                                                                           | run_init_sigma\         |                           |
| *N(0,run_year_sigma).*\                                                                                                                                                                                                                                                                                                                                      | run\[year\]\            |                           |
| The predicted run is\                                                                                                                                                                                                                                                                                                                                        | MLR level 2:\           |                           |
| *r = run + run_year\[year\]*.                                                                                                                                                                                                                                                                                                                                | run_mu\                 |                           |
|                                                                                                                                                                                                                                                                                                                                                              | run_sigma\              |                           |
|                                                                                                                                                                                                                                                                                                                                                              | MLR level 1:\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | run_year_sigma \[year\] |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Mortality.** Survival of juvenile salmon that will enter the ocean (smolts) to become adult return (SAR). Applied simplistically to just the ocean entry year of a salmon, irrespective of freshwater age and duration of marine life. Varies by year, uniform random from mort_min to mort_max.                                                           | mort_min\               | $m_y$                     |
|                                                                                                                                                                                                                                                                                                                                                              | mort_max\               |                           |
|                                                                                                                                                                                                                                                                                                                                                              | mort \[year\]           |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Presence.** The number of salmon on the spawning ground, available to be observed. Derived from the temporal pattern of salmon entering, then residence time before leaving (death or emigration), and the abundance of spawners each year. There are three version of this: simulated, the simulated observed with error, and prediction of the observed. | y_true \[year,day\]\    | $Y_{y,d}$                 |
|                                                                                                                                                                                                                                                                                                                                                              | y_obs \[year,day\]\     |                           |
|                                                                                                                                                                                                                                                                                                                                                              | y_pred \[year,day\]     |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Entry Timing**. *Multi-level.* The temporal pattern of spawners entering the spawning ground.\                                                                                                                                                                                                                                                             | Simulation:\            |                           |
| Level 2, the background, is timing_mu and spread_mu. These are used generate vectors of timing\[year\] and spread\[year\] that are used to simulate spawner presence. The prior knowledge for the level 2 PDD for those is timing_sigma and spread_sigma.\                                                                                                   | timing\[year\]\         |                           |
| For Level 1, the prior knowledge for year effects is timing_year_sigma and spread_year_sigma.                                                                                                                                                                                                                                                                | MLR level 2:\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | timing_mu\              |                           |
|                                                                                                                                                                                                                                                                                                                                                              | timing_sigma\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | MLR level 1:\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | timing_year_sigma       |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Entry Spread**. *Multi-level.*                                                                                                                                                                                                                                                                                                                             | Simulation:\            |                           |
|                                                                                                                                                                                                                                                                                                                                                              | spread \[year\]\        |                           |
|                                                                                                                                                                                                                                                                                                                                                              | MLR level 2:\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | spread_mu\              |                           |
|                                                                                                                                                                                                                                                                                                                                                              | spread_sigma\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | MLR level 1:\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | spread_year_sigma       |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Residence**. Duration, in days, of a spawner on the spawning ground (observation place). Initially constant across years and days within years. This is known to vary by fish sex (males stay longer), and timing (early arrivals stay longer). Likely also varies by fish size (age) and by preceding and present habitat conditions.                     | Simulation:\            | $\lambda$                 |
|                                                                                                                                                                                                                                                                                                                                                              | residence\              |                           |
|                                                                                                                                                                                                                                                                                                                                                              | MLR level 1 prior:\     |                           |
|                                                                                                                                                                                                                                                                                                                                                              | residence_mu\           |                           |
|                                                                                                                                                                                                                                                                                                                                                              | residence_sigma         |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Day.** Day of the solar year, ecological year.\                                                                                                                                                                                                                                                                                                            | day                     | $d$                       |
| Day 0 is winter solstice, Medieval: December 21. January 1: Day 10, Spring equinox: 90, Summer solstice: 181, Fall equinox: 275.\                                                                                                                                                                                                                            |                         |                           |
| <https://www.timeanddate.com/calendar/seasons.html>\                                                                                                                                                                                                                                                                                                         |                         |                           |
| prev_days = 0 31 59 90 120 151 181 212 243 273 304 334\                                                                                                                                                                                                                                                                                                      |                         |                           |
| June 20 = prev_days\[6\]+ 20 +10 = 181\                                                                                                                                                                                                                                                                                                                      |                         |                           |
| Earth orbits Sol in an ellipse: Spring to Fall is \~ 5 days longer than Fall to Spring. Equinoxes and solstices are not 365.256363/4 days apart, not solar days 91.3125, 182.6250, 273.9375                                                                                                                                                                  |                         |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Observation**. The days, each year, when a human (or ?) counted spawners.\                                                                                                                                                                                                                                                                                 | obs\[day, year\]\       | $O_{d,y}$                 |
| Present schedule: random number of observations at random dates each year; constrained to avoid sequential days.\                                                                                                                                                                                                                                            | n_obs_max\              |                           |
| Alternative: Start on the same day every year (Fall equinox + 10 days?) and repeat as Monday, Wednesday, Friday (14 obs/month):\                                                                                                                                                                                                                             | n_obs_min\              |                           |
| 285+ c(0,2,4,7,9,11,14,16,18,21,23,25,28,30)\                                                                                                                                                                                                                                                                                                                | day_min\                |                           |
| = 285 287 289 292 294 296 299 301 303 306 308 310 313 315.\                                                                                                                                                                                                                                                                                                  | day_max\                |                           |
| or Monday, Thursday (10 obs/month):\                                                                                                                                                                                                                                                                                                                         | n_days                  |                           |
| 285+ c(0,3,7,10,14,17,21,24,28,31)\                                                                                                                                                                                                                                                                                                                          |                         |                           |
| = 285 288 292 295 299 302 306 309 313 316.\                                                                                                                                                                                                                                                                                                                  |                         |                           |
| The display of sampling involves an x-axis range, *day_min* to *day_max*, a total of *n_days*; calculated from entry parameters.                                                                                                                                                                                                                             |                         |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Sampling Error**. The precision of an observation is Poisson (variance = mean) or neg.binomial). Results are non-negative integers and sampling error is proportional to spawner abundance (multiplicative, error is a percent). Then an observation of an abundance is $O_{y,d}=rlnorm(log(A_{y,d}), \sigma_O)$ where $\sigma_O\approx 0.15$.\            | sigma_obs               | $\sigma_O$                |
| qlnorm(c(0.025,0.25,0.75,0.975),log(1e4), 0.25)\                                                                                                                                                                                                                                                                                                             |                         |                           |
| 6,126 8,448 11,836 16,323 \# 95% and 50% CIs                                                                                                                                                                                                                                                                                                                 |                         |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Productivity**. The ratio of juveniles to spawners, on the order of 70. The juveniles become returning adults after 1 or 2 years freshwater, and 1 to 3 years saltwater. This is a fixed value in the present model. This model adds a lognormal random component to mean productivity.                                                                    | prod,\                  | $P, \sigma_P, P_y$        |
|                                                                                                                                                                                                                                                                                                                                                              | prod_sigma,\            |                           |
|                                                                                                                                                                                                                                                                                                                                                              | prod_year               |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
| **Life History.** Apportions age at return (fresh then salt). These do not vary by year. Organized as (1) hist_fresh: probabilities for smolts being 1 or 2 lake years (1.x, 2.x) so a fraction emigrate later and encounter different SAR, then (2) hist_salt: 1, 2, or 3 marine years. Example:\                                                           | hist_fresh,\            | $h_f,h_s$                 |
| LH for ages (1.1,1.2,1.3), (2.1,2.2,2.3)\                                                                                                                                                                                                                                                                                                                    | hist_salt               |                           |
| = hist_fresh\*hist_salt\                                                                                                                                                                                                                                                                                                                                     |                         |                           |
| = c(0.8\*c(0.2,0.6,0.2), 0.2\*c(0.2,0.6,0.2))\                                                                                                                                                                                                                                                                                                               |                         |                           |
| = (0.6 0.48 0.16), (0.04 0.12 0.04). Sum = 1.                                                                                                                                                                                                                                                                                                                |                         |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+
|                                                                                                                                                                                                                                                                                                                                                              |                         |                           |
+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+-------------------------+---------------------------+

: Table 1. Parameters for simulated observations of spawner abundance.

The format for the data input to the Stan model is new. There is a different number of observations each year. The dates for those observations vary each year. The count $y_{true}$ of live spawners, and the day of the year $day$ of that count, are in two ragged arrays: $day[n\_year,max\_obs]$and $y_{true} [n\_year,max\_obs]$,\
where $max\_obs$ is the maximum number observations in any of the $n\_year$ years. The number of observations each year (the lengths of valid data in each row of the ragged array) is a vector $n\_obs\_year[n\_year]$, so the data in year $i$ is $y[i, 1:n\_obs\_year[i]]$.

The model for error structure of the observations is critical to understanding how model predictions fit observations. In the simulated data, sampling error is a fixed percentage of the observation, one might say an observations was plus or minus 20%. Thus the error is multiplicative not additive. $y_{obs}=y\times \mathcal{N}(1,\sigma)$ not $y_{obs}=y + \mathcal{N}(0,\sigma)$. In that case an observation $y_{obs}$ of 100 spawners $y_{true}$ has a 68% chance of being between 80 to 120, $\mu \times(\pm \ \mathcal{N}(1,\sigma))$, and similarly for an observation of 100,000: 80,000 to 120,000. This is the same as saying the sampling distribution for an observation is lognormal.

````{=tex}
```{=tex}
\begin{equation*} \begin{aligned}
y_{obs} &=\text{rlnorm}(\text{log}(y),\sigma) \\
 &= (1/y) e^{\mathcal{N}(\text{log}(y),\sigma)}
\end{aligned} \end{equation*}
```
````

This is coded as

| n = n_obs_year\[year\];
| y_obs\[1:n\] \~ lognormal(log(y_pred\[year,1:n\]), sigma); // relative error

where *y_pred* is a ragged array.

# Simulated Data

Derived from *Simulate Spawners.qmd*, but number and dates for observations vary by year. Just one river.

## Set Parameters

Starting with four years of spawner abundance, subsequent abundances are derived from $prod\equiv$ **juveniles per spawner** (plus variability), apportioning juveniles to life history fractions, and $mort\equiv$ **survival of smolts to adult returns** SAR. The initial years are random lognormal: $\text{rlnorm}(4, \text{log}(1e4), 1)$ .

```{r}
#| label: constants
# Set constants for simulation. run, timing, spread.
n_years = 20     # N rows in data
n_obs_max = 10   # upper limit number of samples each year. N columns in data.
n_obs_min = 4    # lower limit number of samples each year
mort_min = 0.01  # uniform random survival from juv to spawner
mort_max = 0.15  # uniform random survival from juv to spawner
prod = 1/((mort_min + mort_max) /2)  # juveniles per spawner, inverse of mean mortality.
run_init_mu = 1e4;     # mean for initial 4 years spawner run, Gaussian
run_init_sigma = 10000;  # sd to initialize
run_mu = 1e4;            # prior knowledge of PDD runs
run_sigma = 5e3;         # prior knowledge of PDD runs
LH_fresh = 0.8           # frac smolts 1.x;frac 2.x=1-LH_fresh
LH_salt = c(0.2,0.6,0.2) # frac sea years x.1, x.2, x.3 
# Spawner entry every year.
# These are also level 2 parameters in Stan model.
timing_mu = 300   # mean day spawners enter spawning grounds
timing_sigma = 5
timing_year_sigma = 5  # year effect on. range ~20 days: 290,310.
# mean over all years for spread of spawners entering in a year.
spread_mu = 7 # mean of spread in days for spawners entering. 95% within 28 days
spread_sigma = 1
spread_year_sigma = 1 # year effect on spread.
residence = 12            # days on spawning ground
residence_mu = 12         # prior
residence_sigma = 3       # prior
sigma_sampling = 0.15     # sampling error, proportional

# check max range of spawning
extremes = c(timing_mu - 2*timing_sigma - 
               (spread_mu + 2*spread_sigma),
             timing_mu + 2*timing_sigma + 
               (spread_mu + 2*spread_sigma) )
cat('extreme days:', extremes[1],'to',extremes[2],'range:',extremes[2]-extremes[1],'\n')

cat("\n Scalar Parameters:", 
    "\n n_years:", n_years, '  n_obs_min:', n_obs_min, '  n_obs_max:',n_obs_max,
    "\n prod:", prod, 
    "\n  mort_min:", mort_min, "  mort_max:", mort_max,
    "\n run_init_mu:",run_init_mu,  "  run_init_sigma:",run_init_sigma,
    "\n LH_fresh:",LH_fresh, "  LH_salt:",LH_salt,
    "\n timing_mu:",timing_mu,    "  timing_sigma, year effect:", timing_sigma,
    "\n spread_mu:",spread_mu,  "  spread_sigma, year effect:", spread_sigma,
    "\n residence:", residence,
      "  residence_mu:", residence_mu,"  residence_sigma:", residence_sigma,
    "\n sigma_prop:", sigma_sampling, '\n')
 
```

### Apply Parameters

Derive resulting constants and inputs to simulation model.

```{r}
#| label: apply
# apply constants  
Trunc_rnorm = function (n, m=0, s=1, lowest=-Inf, highest=Inf){
    a = numeric(n)  # zeros
    j = rep(TRUE, n)  # which elements in a to replace, start with all.
    while(sum(j) > 0){
    a = c(a[!j], rnorm(sum(j), m, s) ) # replace with random draws
    j= a < lowest | a > highest # find those out of range
    }
    return(a)
   # Trunc_rnorm(n=10, m=10, s=5, lowest=0) # test 
}

mort =   runif(n_years, mort_min, mort_max ) # SAR, entry year only
timing = rnorm(n_years, timing_mu, timing_sigma) %>% round(0) 
spread = rnorm(n_years, spread_mu, spread_sigma) %>% round(1) 
day_min = min(timing) - 2* max(spread) %>% as.integer()
day_max = max(timing) + 2* max(spread) %>% as.integer()
n_days = 1+ day_max - day_min
run = numeric(n_years) 
# initial 4 years of spawner runs, needed to start popn dynamics
run[1:4] = Trunc_rnorm(n=4, run_init_mu, run_init_mu, lowest=1000) %>% round(0) 
#run[1:4] = rlnorm(4, log(run_init_mu), 2) %>% round(0) # start for sim
cat("\n Patterns From Scalar Parameters:", 
    '\n day_min:', day_min, '  day_max:', day_max, '  n_days:',n_days, 
    '\n mort each year: \n',    round(mort,2),
    '\n timing each year: \n',  timing, 
    '\n spread each year: \n',  spread, 
    '\n initial runs: \n', run[1:4], '\n')
```

## Salmon Population Dynamics

Generate *n_years* of spawner abundance, starting with 4 years random values from parameters. The age composition of spawners is available from this, possibly for spawner residence time as a function of size or age. Survival from smolts to spawners is simplistic in the version.

```{r}
#| label: pop_dyn
# a function to generate spawner abund each year.
Pop_dyn = function(spawners, LH_fresh, LH_salt, prod, mort){
# length of spawners sets length of result. NA after first 4 years. 
# abundance is total spawner population abundance, ages merged. 
# calculate returns from spawners for an additional 3 years. 
    # if 20 years of spawners, year 17 is last addition to year via returns age 1.1 
    # but year 17 calculations produce returns age 2.3 for year 23. 
    # truncate before return    
    n_years= length(spawners)
    abundance = numeric(n_years+3) 
# transfer starting values for spawners
    abundance[1:4]= spawners [1:4] 
# abundance[4] is increased by returns 1.1 from spawners[1]
# if n_year = 20, last spawners useful is year 17
    for (y in 1:(n_years-3)) {  
        juv=abundance[y]*prod 
# apportion to freshwater duration, one or two years
# assign mortality according to ocean entry year
        age_1x = juv *    LH_fresh  * mort[y+2]
        age_2x = juv * (1-LH_fresh) * mort[y+3]
# apportion to saltwater duration, one, two, or three years
    # age_11 = age_1x * LH_salt[1]
    # age_12 = age_1x * LH_salt[2]
    # age_13 = age_1x * LH_salt[3]
    # age_21 = age_2x * LH_salt[1]
    # age_22 = age_2x * LH_salt[2]
    # age_23 = age_2x * LH_salt[3]
        abundance[y+3] %<>% + (age_1x * LH_salt[1])  # age_11  
        abundance[y+4] %<>% + (age_1x * LH_salt[2] + age_2x * LH_salt[1]) # 1.2+2.1
        abundance[y+5] %<>% + (age_1x * LH_salt[3] + age_2x * LH_salt[2]) # 1.3+2.2
        abundance[y+6] %<>% + (age_2x * LH_salt[3])  # age_23
    }
    return(abundance[1:n_years]) # truncated
}
```

## Resulting Population Dynamics

```{r}
#| label: test_pop_dyn
 run = Pop_dyn(run, LH_fresh, LH_salt, prod, mort)
y_max=1.1*max(run)

plot_dat= data.frame(Year=1:n_years, Spawners = run)
ggplot(data=plot_dat, aes(Year, Spawners)) + theme_SAA() +
    coord_cartesian(xlim=c(0,n_years+1), ylim=c(0,y_max), expand=FALSE) +
    geom_line() + geom_point()

```

## Spawner observations with sampling error

### Spawner Presence

Spawner *run* each year is converted to spawner *presence* each day on the spawning grounds as a cumulative normal distribution of entries by day and corresponding cumulative exits after *residence* days later. These parameters vary by year with the exception of residence

Typically Gaussian, $G(\mu_{E,y}, \sigma_E)$, with the mean day of entry varying by year but the spread of spawner presence is constant across years. So $T_y=\mu_{E,y} =G(\mu_T, \sigma_T)$ where $\mu_T$ is the mean day across years of the mean day of spawner entry. $E_{y,d}=G(T_y, \sigma_E)$.

First, the model of spawner abundance for observations days. It is not necessary to determine the integral of the difference of two cumulative probability distributions. This is simply $A_t=\mathcal{N}(t,\mu,\sigma) - \mathcal{N}(t,\mu+\tau,\sigma))$ where $\tau$ is the lag, the residence time, between a spawner entering and exiting. At any instant *t*, the number of spawner present is $A_t$, the basis for observations. For realistic values of $\sigma \text{ and }  \tau, \ A_t$ is a normal distribution. The sum of this abundance each day will be the number of spawners entering multiplied by the residence time. This in in accordance with the calculation of AUC, where the cumulative abundance of live spawners (by linear interpolation) has to be corrected by the residence time.

```{r}
#| label: spawner_presence
Presence = function(days, mu, sigma, residence){
    # pnorm is cumulative probability, 
    # pnorm( c(-1.96,1.99))= 0.025, 0.975
    a = (pnorm(days,mu,sigma) - pnorm(days,mu+residence,sigma))
    return(a)
}
#presence= 100* Presence( days=c(14,17,20,23,26,29,32), mu=20, sigma=4, residence= 6); round(presence,0)
#  7 21 43 55 43 21  7
# sum(presence*3)/6
#  98.64256  # this is AUC (without bracketing days of zero fish)
```

### Irregular Sampling Days

Apply the Presence() function and select values on sampling days, then observations with multiplicative sampling error of those values.\
Three functions are required:

1.  **how many observations?** For each year, a random uniform integer between *n_obs_min* and *n_obs_max*, say 3 to t0.

2.  **Draw_days()** generates the dates for that number of observations each year, within a bounded range, *day_min* to *day_max*. These are is random draws from a truncated normal distribution: a *while loop* replaces values out of range.

3.  **Sampling()** applies the preceding functions, but replaces consecutive days. That is done by calling Draw_days() for replacement of trial values.

```{r}
#| label: f_sampling
#| 
  Draw_days = function (n, timing_mu, spread_mu, widen=1, day_min, day_max){
    a = rnorm(n, timing_mu, spread_mu*widen) %>% 
        sort  %>% round(0);  # cat('a1    ',a,'\n')
    # replace draws outside day_min to day_max   
    while ( ((min(a) < day_min) | (max(a) > day_max)) == TRUE){
        j=which( (a < day_min) | (a > day_max) )
        a = c(a[-j], rnorm(length(j), timing_mu, spread_mu*widen)) %>% 
        sort  %>% round(0);  # cat('a2    ',a,'\n')
    }
  return(a)
  }

Sampling = function(n_obs_min,n_obs_max,day_min, day_max, timing_mu, spread_mu ){
# generate number of samples, random uniform.
# generate dates for these samples within a bounded range of days.
# these days cannot be consecutive. min gap is 1 days.

# get number of days 
  n = runif(1,n_obs_min,n_obs_max) %>% round(0) %>% as.integer
# get dates  
  a = Draw_days(n, timing_mu, spread_mu, widen=2, day_min, day_max) 
  # cat('a final:',a,'\n')
# elimate same or consecutive days
  while (1 >= min(a[2:n] - a[1:(n-1)]) ) { # if same or consecutive days
    # find them
    j= which ( 1 >= (a[2:n] - a[1:(n-1)]) ) 
    # create replacements
    b =  Draw_days(length (j), timing_mu, spread_mu, widen=1, day_min, day_max)    
    a = c(a[-(j+1)], b) %>% sort # goes to test again in while
  }    
  return(a)
}

```

Applying *Sampling()* for observations dates is the set-up for *Presence()* to provide *y_true*, the number of spawners available on each observation day. Then the true number is subject to proportional sampling error (like 15% accuracy), applied as a random lognormal draw to generate the recorded observation: $y_{obs} = \text{rlnorm}(n,\text{log}(y_{true}), \sigma)$.

The results are stored:

-   **n_obs_year:** a vector with the number of days sampled each year.
-   **day**: a matrix where each row is the dates of observations for a year. There are n_obs_year\[year\] valid observations, and the remainder of the *n_obs_max* columns are NA. This is a *ragged array.*
-   **y_true**: a matrix corresponding to *day*.
-   **y_obs**: a matrix corresponding to *day*.

```{r}
#| label: obs
day = y_obs = y_true = matrix(0.0, n_years, n_obs_max)
n_obs_year = numeric(n_years)

for (year in 1:n_years){ 
    days = Sampling(n_obs_min,n_obs_max,day_min,day_max, timing_mu, spread_mu ) # samples in one year. SORTED.
    n_obs_year[year] = n = length(days)
    day[year, 1:n] = days     # remainder of this row in the array is NA
# find the spawner density on those days
    presence = run[year] * Presence(days,timing[year],spread[year],residence);
    y_true[year,1:n ] = round(presence,0) # true count
}

# sampling error is 15%, or at least, sigma in lognormal is 0.15 
for(year in 1:n_years){
    n = n_obs_year[year]
    y_obs[year, 1:n] = rnorm(n, log(y_true[year, 1:n]), sigma_sampling) %>% exp %>% floor # as integer counts for Poisson sampling: y_obs ~ Poisson(y_pred)
}
```

## Plot obs vs true

```{r}
#| label: obs_true
#| warning: false
Ragged_2_Vector = function(a, b){
# a is a ragged matrix or array, data does not fill rows, remainder is NA
# b is the length of data for each row. So b has length nrow(a)
  n_rows = nrow(a) 
  x= numeric(sum(b)) 
  k1=0;k2=0;
  for(j in 1:n_rows){
    k1=k2+1
    k2=k2+b[j]
    x[k1:k2]=a[j,1:b[j] ]
  }
  return(x)
}
Repeated = function(a,b){
    # repeats each element of a according to corresponding element in b
  if(length(a) != length(b)) stop('Error in Repeated: lengths differ')
  x= numeric(sum(b)) 
  k1=0;k2=0;
  for (j in 1:length(a)) {
    k1=k2+1
    k2=k2+b[j]
    x[k1:k2]= rep(a[j], b[j])
  }
  return(x)
}
plot_dat = data.frame(
    year =   Repeated(2001:(2000+n_years), n_obs_year),
    day =    Ragged_2_Vector(day, n_obs_year),
    y_true = Ragged_2_Vector(y_true, n_obs_year),
    y_obs =  Ragged_2_Vector(y_obs, n_obs_year) )
# year day y_true y_obs
# 2001 297   1136  1250 ... for sum(n_obs_year) rows.

theme_SAA()
p1=ggplot(plot_dat, aes(y_true, y_obs) ) + 
    geom_point() + geom_abline(slope=1, intercept=0) +
    labs(title='Linear', x='Spawners',y='Estimates')
p2=ggplot(plot_dat, aes(y_true, y_obs) ) + theme_SAA() +
    geom_point() +
    scale_x_log10() + scale_y_log10() +
    geom_abline(slope=1, intercept=0)+
    labs(title='Logarithmic', x='Spawners',y='')
#  coord_trans(x = "log10", y = "log10") 
p1 + p2 

```

## Plot spawner timing by year

Examine the normal distributions of spawner timing by years. This is the pattern without annual abundance of spawners.

```{r}
#| label: spawner_timing
# generate underlying spawner density for all days all years
# needs year, day, y_true
n = (day_max-day_min)+1  # number of days simulated
all_days=data.frame(
    year=rep( 2001:(2000+n_years), each=n),
    day=rep(day_min:day_max, n_years), 
    y=NA)
k1=0
for(year in 1:n_years){
  k1=k1+1; k2 = (k1+n)-1
  all_days[k1:k2, 3] = run[year]* 
    Presence(day_min:day_max,timing[year],spread[year],residence);
  k1=k2
}

theme_SAA()
y_max = 1.1* max(plot_dat[,3:4])
ggplot(plot_dat, aes(day,y_obs)) + #  theme_SAA() +
    geom_point(size=0.5) +
    geom_line(data=all_days, mapping=aes(day,y), linewidth=0.2) +
    facet_wrap(year~.) +
        labs(x="Day of the Year",y="Live Spawners",title='Simulated Observations') 
ggsave('./images/simulated_observations.png', height=7,width=5.5)

```

## Data and Prior Knowledge for Stan Model

Priors.

```{r}
#| label: priors_4_stan
# data # prior knowledge of level 2 distributions.  14 of these.
#   real run_mu_prior;           // run_mu is a parameter
#   real run_mu_sigma_prior;     // with a distribution
#   real run_sigma_prior;        // run_sigma is a parameter
#   real run_sigma_sigma_prior;  // with a distribution
#   real timing_mu_prior;
#   real timing_mu_sigma_prior;
#   real timing_sigma_prior;
#   real timing_sigma_sigma_prior;
#   real spread_mu_prior; 
#   real spread_mu_sigma_prior;
#   real spread_sigma_prior; 
#   real spread_sigma_sigma_prior;
#   real residence_mu_prior;
#   real residence_sigma_prior;

# model level 2
  # run_mu       ~ normal(run_mu_prior, run_mu_sigma_prior); 
  # run_sigma    ~ normal(run_sigma_prior, run_sigma_sigma_prior; 
run_mu_prior =             median(run) # param. about 20,000
run_mu_sigma_prior =       0.25 * run_mu_prior
run_sigma_prior =          0.25 * run_mu_prior # param for sd of runs
run_sigma_sigma_prior =    0.25 * run_sigma_prior
  # timing_mu    ~ normal(timing_mu_prior, timing_sigma_prior);
  # timing_sigma ~ normal(timing_sigma_prior, timing_sigma_sigma_prior;
timing_mu_prior =          timing_mu # param. 300 from sim
timing_mu_sigma_prior =    timing_sigma # 5 from sim 
timing_sigma_prior =       timing_sigma # param. 5 from sim 
timing_sigma_sigma_prior = 1 
# spread_mu    ~ normal(spread_mu_prior, spread_sigma_prior);
# spread_sigma ~ normal(spread_sigma_prior, spread_sigma_sigma_prior);
spread_mu_prior =          spread_mu # param. 7 from sim
spread_mu_sigma_prior =    spread_sigma # 1 from sim
spread_sigma_prior =       spread_sigma # param. 1 from sim
spread_sigma_sigma_prior = 0.25 * spread_sigma_prior
# residence    ~ normal(residence_mu_prior, residence_sigma_prior);
## this is not a multilevel parameter
residence_mu_prior = residence_mu # param. 12 from sim
residence_sigma_prior = residence_sigma 

cat('\nPrior knowledge of parameter distributions\n'); 
cat(
  '\n run_mu_prior:', run_mu_prior, 'run_mu_sigma_prior:',run_mu_sigma_prior,
  '\n run_sigma_prior:', run_sigma_prior, 'run_sigma_sigma_prior:',run_sigma_sigma_prior,
  '\n timing_mu_prior:',timing_mu_prior,  'timing_mu_sigma_prior:', timing_mu_sigma_prior, 
  '\n timing_sigma_prior:', timing_sigma_prior, 'timing_sigma_sigma_prior:', timing_sigma_sigma_prior,
  '\n spread_mu_prior:',spread_mu_prior,  'spread_mu_sigma_prior:',spread_mu_sigma_prior, 
  '\n spread_sigma_prior:',spread_sigma_prior, 'spread_sigma_sigma_prior:', spread_sigma_sigma_prior,
  '\n residence_mu_prior:',residence_mu_prior,'residence_sigma_prior:', residence_sigma_prior
)

```

Data

```{r}
#| label: dat_4_stan

#data 
#  int n_years;                    // rows in data
#  int n_max_obs;                  // columns in data
#  int n_obs_year [n_years];       // how many obs each year
#  matrix [n_years, n_max_obs] day;        // dates,    ragged, reals 
#  array  [n_years, n_max_obs] int y_obs;  // observed, ragged, integers
dat = list(
  run_mu_prior = run_mu_prior, 
  run_mu_sigma_prior = run_mu_sigma_prior,
  run_sigma_prior = run_sigma_prior,
  run_sigma_sigma_prior = run_sigma_sigma_prior,
  timing_mu_prior = timing_mu_prior, 
  timing_mu_sigma_prior = timing_mu_sigma_prior, 
  timing_sigma_prior = timing_sigma_prior,
  timing_sigma_sigma_prior = timing_sigma_sigma_prior,
  spread_mu_prior = spread_mu_prior, 
  spread_mu_sigma_prior = spread_mu_sigma_prior,
  spread_sigma_prior = spread_sigma_prior,
  spread_sigma_sigma_prior = spread_sigma_sigma_prior,
  residence_mu_prior =residence_mu_prior, 
  residence_sigma_prior = residence_sigma_prior,
    n_years=n_years, # data
    n_max_obs=n_obs_max,
    n_obs_year=n_obs_year,
    day=day,
    y_obs=y_obs)
cat('\n dat for Stan:\n'); str(dat)

```

## Stan Code

The model is a character string. This is checked for syntax by function **stanc().** Upon passing those checks, the result is C+ code. That result is compiled to be part of a Stan model by function **stan_model().** Applying the model to data, "fitting" the model, is a subsequent function **sampling().**

```{r}
#|label: multi_year_gauss
#|cache: true
MultiYearGauss_code <- ' 
functions {  
vector Presence(int n, row_vector days, real abund, 
                real mu, real sigma, real lag) {
    vector [n] presence; 
    for (j in 1:n){
        presence[j] = normal_cdf(days[j], mu,    sigma) - 
                      normal_cdf(days[j], mu+lag,sigma);
    }
    return abund .* presence;
  }
}

data {
//prior knowledge of level 2 distributions. 14 of these.
  real run_mu_prior;           // run_mu is a parameter
  real run_mu_sigma_prior;     // with a distribution
  real run_sigma_prior;        // run_sigma is a parameter
  real run_sigma_sigma_prior;  // with a distribution
  real timing_mu_prior;
  real timing_mu_sigma_prior;
  real timing_sigma_prior;
  real timing_sigma_sigma_prior;
  real spread_mu_prior; 
  real spread_mu_sigma_prior;
  real spread_sigma_prior; 
  real spread_sigma_sigma_prior;
  real residence_mu_prior;
  real residence_sigma_prior;
//data  
  int n_years;                            // rows in data
  int n_max_obs;                          // columns in data
  array  [n_years] int n_obs_year;        // number of obs each year
  matrix [n_years, n_max_obs] day;        // dates, ragged, reals 
  array  [n_years, n_max_obs] int y_obs;  // obs,   ragged, integers
}
parameters { 
//---- level 2: overall ----
  real <lower=0> run_mu;        // mean all years
  real <lower=0> run_sigma;     // sd   all years
  real <lower=0> timing_mu;     // mean all years
  real <lower=0> timing_sigma;  // sd   all years
  real <lower=0> spread_mu;     // mean all years
  real <lower=0> spread_sigma;  // sd   all years
  real <lower=0> residence;     // duration observable, no year effect
//---- level 1: instance ----
  vector  [n_years]  run_year;     // run    each year
  vector  [n_years]  timing_year;  // timing each year
  vector  [n_years]  spread_year;  // spread each year
}
model {
//---- local variables, values within a year ----
  vector [n_max_obs] y_pred;  // predicted obs 
  real  r;                    // predicted run
  real  t;                    // predicted timing
  real  s;                    // predicted spread
  int   n;                    // how many obs in this year

//---- prior knowledge of level 2 parameter distributions ----
  run_mu       ~ normal(run_mu_prior,       run_mu_sigma_prior); 
  run_sigma    ~ normal(run_sigma_prior,    run_sigma_sigma_prior); 
  timing_mu    ~ normal(timing_mu_prior,    timing_sigma_prior);
  timing_sigma ~ normal(timing_sigma_prior, timing_sigma_sigma_prior);
  spread_mu    ~ normal(spread_mu_prior,    spread_sigma_prior);
  spread_sigma ~ normal(spread_sigma_prior, spread_sigma_sigma_prior);
  residence    ~ normal(residence_mu_prior, residence_sigma_prior); // no year effect

//---- level 1 parameter distributions sampled from level 2 ----
  run_year     ~ normal(run_mu,run_sigma);
  timing_year  ~ normal(timing_mu,timing_sigma);
  spread_year  ~ normal(spread_mu,spread_sigma);

//---- predict from sampled parameters ---- 
//print(" run_mu: ", run_mu, " run_sigma: ", run_sigma,
//  " timing_mu: ", timing_mu, " timing_sigma: ", timing_sigma,
//  " spread_mu: ", spread_mu, " spread_sigma: ", spread_sigma,
//  " residence ", residence) ; 
//print(" run_year: ", run_year);
//print(" timing_year: ", timing_year);
//print(" spread_year: ", spread_year);
    
  for (year in 1:n_years) {
    r = run_year[year]; 
    t = timing_year[year];
    s = spread_year[year];
    n = n_obs_year[year]; 
    y_pred[1:n] = Presence(n,day[year],r,t,s,residence);
//print(y_obs [year, 1:n]);
//print(y_pred[1:n]);

//---- accumulate probability of data given prediction ----
    target += poisson_lupmf(y_obs[year, 1:n] | y_pred[1:n]); # y_obs is integers
  }    
} '

stanc_ret <- stanc(model_code=MultiYearGauss_code) # check fussy syntax, compile 
MultiYearGauss <- stan_model (stanc_ret=stanc_ret, model_name = 'MultiYearGauss') # create model
```

Expose Function: Presence()

# Fit Model to Observations

*No prediction, no science.*

$y_{obs}$ is from imprecise sampling of simulated $y_{true}$, and applied to challenge predictions, $y_{pred}$, from this (of all possible) model.

```{r}
#|label: sampling_MultiYearGauss
#|cache: true
init_1_chain = list( 
  list( 
    run_mu= 20000, run_sigma=2000, timing_mu = 300, timing_sigma = 5, 
    spread_mu =7, spread_sigma = 1, residence=12,
    run_year =    rnorm(20, 20000, 2000), 
    timing_year = rnorm(20,   300, 5),
    spread_year = rnorm(20,    12, 2) 
  )
)

init_4_chains = list(
  list(
    run_mu=22000, run_sigma=4000, timing_mu= 310, timing_sigma=7, 
    spread_mu=9, spread_sigma=2, residence = 13,
    run_year =    rnorm(20, 20000, 4000), 
    timing_year = rnorm(20, 300, 5),
    spread_year = rnorm(20, 12, 2) 
  ),
  list(
    run_mu=21000, run_sigma=3500, timing_mu=305, timing_sigma=6, 
    spread_mu=8, spread_sigma=2, residence=13,
    run_year =    rnorm(20, 20000,4000), 
    timing_year = rnorm(20, 300, 5),
    spread_year = rnorm(20,  12, 2) 
  ),
  list(
    run_mu=20000, run_sigma=3000, timing_mu=300, timing_sigma=5, 
    spread_mu=7, spread_sigma=2, residence=13,
    run_year =    rnorm(20, 20000,4000), 
    timing_year = rnorm(20, 300, 5),
    spread_year = rnorm(20,  12, 2) 
  ),
  list(
    run_mu=19000, run_sigma=2500, timing_mu=295, timing_sigma=4, 
    spread_mu=6, spread_sigma=2, residence=13,
    run_year =    rnorm(20, 20000,4000), 
    timing_year = rnorm(20, 300, 5),
    spread_year = rnorm(20,  12, 2) 
  )
)

#s1 <- sampling( MultiYearGauss, data=dat, init=init_1_chain,chains=1, iter=2)
s1 <- sampling( MultiYearGauss, data=dat, init=init_4_chains,
                chains=4, iter=10000, warmup=2000, cores=4, refresh=2000)
```

### Marginal Distributions

```{r}
#|label: prt_sampling_s1
print (s1)
plot(s1, pars=c('run_year'))
plot(s1, pars=c('timing_year'))
#plot(s1, pars=c(''))
```

### Compare fitted parameters to original

```{r}
#| label: compare

a=s1@.MISC[["summary"]][["msd"]]  # mean and sd from fit
a1= numeric(dim(a)[1])
a1[ 1:7] = c(mean(run), sd(run),mean(timing), sd(timing),mean(spread), sd(spread), residence)
a1[ 8:27] = run
a1[28:47] = timing
a1[48:67] = spread
a1[68] = NA
a = cbind(a, a1) %>% round(2); rm(a1);
colnames(a) = c("Predicted","Stdev","Observed")
print(a)
write.csv(a, file='./results/fit and obs poisson residuals 2024-08-31.txt')
# plot pred vs true
a1 = a[-c(1:7, 68),] %>% as.data.frame # drop scalars and lp__
a1$param = rep(c('Run','Timing','Spread'), each = 20)
ggplot(a1, aes(Observed, Predicted)) + 
    facet_wrap(param~., scales='free') +
    geom_point() + geom_abline(slope=1, intercept=0) 

ggsave('./results/pred_obs_poisson_2024_08_31.pdf', width=5.5, height=5.5) 
```

### Save Results

```{r}
#| label: stash
#| eval: false
# saveRDS( obs_sampl, './results/obs_sampl.Rdata')
# saveRDS( obs_true, './results/obs_true.Rdata')
```
